### 2019-06-10 paper/main.tex
main

\documentclass[twocolumn]{aastex62}
\submitjournal{ApJL}
\title{Our peculiar motion across the Universe}
\shortauthors{Forero-Romero}

\usepackage{import}
\usepackage{hyperref}
\usepackage[T1]{fontenc}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage{float}

\begin{document}
\title{Our peculiar motion across the Universe}

\author{Jaime E. Forero-Romero}
\affiliation{Departamento de F\'isica\\
Universidad de los Andes\\
Cra. 1 No. 18A-10\\
Bogot\'a, Colombia}

\begin{abstract}
We
\end{abstract}

\keywords{Astronomy databases (83), Transient detection (1957),
  Astrostatistics tools (1887)}   
\section{Introduction}

\cite{2009PhRvD..80d3005E}

\bibliographystyle{aasjournal}
\bibliography{references}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.9\textwidth]{cumulative_probability_18.pdf}
\end{center}
\caption{A}
\label{fig:proba}
\end{figure*}

\end{document}

Large scale automatic detection and classification of astronomical
transients is happening within surveys such as Pan-STARRS1
\citep{2004SPIE.5489...11K}, the Palomar Transient Factory
\citep{2009PASP..121.1395L},  the Catalina Real-Time Transient Survey
\citep{2009ApJ...696..870D}, the All-Sky Automated Survey for
SuperNovae \citep{2014ApJ...788...48S}, the Zwicky Transient
Factory \citep{2019PASP..131a8002B}.
Besides the large amount of data, transient
classification is hard because the data is usually heterogeneous,
unbalanced, sparse, unevenly sampled and with missing information.

These two characteristics (size and heterogeneity) have
motivated the application of Machine Learning (ML) algorithms to face
this challenge.  
For instance, Random Forests, MultiLayer Perceptron and K-Nearest
Neighbours have been used on lightcurves to classify transients from
the Catalina Real Time Transient Survey \citep{1601.03931};  
convolutional neural networks have been used as 
input to automatic vetting algorithms (quick classification of bogus
vs. real transients) based on data from the SkyMapper Supernova and
Transient  Survey and the High cadence Transient Survey (HiTS)
\citep{1708.08947,1701.00458}.

The success of these examples, and any other ML implementation,
depends on the quality of the training dataset.
New ML results usually come from groups internal to an observational
collaboration because they have the internal know-how (and, sometimes,
privileged access) to build training datasets.
This difference in data access makes it challenging for the broader
astronomical community to rebuild a training dataset, perform comparisons with
published results and suggest new algorithms.

Nonetheless, building traning datasets has been facilitated by the
publication of large databases from different observational projects.
Other collaborations have directly published large datasets of simulated
hoping to trigger more involvement from the ML in astronomy
community at large to develop new classification algorithms \citep{2018arXiv181000001T}.

In this paper we aim at bridging the data access gap.
To this end we compile and publish in easy-to-access files a dataset
that can be used to train and test different ML algorithms for
transient classification. 
We use public data from the Catalina Real-Time Transient Survey
(CRTS) \citep{1111.2566}, an astronomical survey searching transient
and highly variable objects as base for the dataset.
Here, in the first paper, we present the lightcurve data.
In a second paper we will present a curated imaging dataset from the same
survey.

This paper is structured as follows.
In Section \ref{sec:data} we present the CRTS and the steps we follow
to build the dataset.
Then, in Section \ref{sec:repository} we describe its main features together
with the repository structure gathering the files and Python code to explore it. 
In Section \ref{sec:ml_tests} we show how this dataset can be used to 
perform tests using ML methods following a similar approach as \cite{1601.03931},
and the experiments that we perform. 
We finalize in Section \ref{sec:conclusions} with a summary of the
main features of our dataset and the results of our experiments.

\section{The lightcurve dataset} 
\label{sec:data}

We use public data from the Catalina Real-Time Transient Survey
(CRTS) \citep{2009ApJ...696..870D}, an astronomical survey searching transient
and highly variable objects.
The CRTS covered 33000 squared degrees of sky and took data since 2007. 
Three telescopes were used: Mt. Lemmon Survey (MLS), Catalina Sky 
Survey (CSS), and Siding Spring Survey (SSS). So far, CRTS has 
discovered more than $15000$ transient events.
We use data from the CSS telescope, which is an f/1.8 Schmidt
telescope located in the Santa Catalina Mountains in Arizona.
The telescope is equipped with a 111-megapixel  detector, and covered
4000 square degrees per night, with a limiting magnitude of 19.5 in
the V band.

Putting together the lightcurves for MANTRA implies
cross-matching different files in the legacy CRTS webpage:
\url{http://nesssi.cacr.caltech.edu/DataRelease/CRTS-I_transients.html}. 
The photometry is stored in two different kinds of files: \verb"phot"
that come from the main photometry database and \verb"orphan" that
correspond to transients not associated with the 500 million sources
in the main photometry database.
There are also \verb"out" files that must be used to link transient
IDs to database IDs.

For each one of the 5540 transients reported and classified in the
archival webpage \url{http://nesssi.cacr.caltech.edu/catalina/All.arch.html} we use its
transient IDs and its database IDs to look for the lightcurves in the
\texttt{phot} and \texttt{orphan} files. 
Only 4982 transients can be linked to available data to reconstruct
their lightcurves. 
Furthermore, some of these lightcurves are duplicated, i.e. they had
the same number of observations, Modified Julian Date (MJD) and magnitude measurement. 
We ignore the duplicates to end up with 4869 unique transients with an 
associated lightcurve. Figure \ref{fig:transients} summarizes this process.

The CRTS dataset already provides a classification. 
The most numerous classes are: supernovae,
cataclysmic variable stars, blazars, flares, asteroids, active
galactic nuclei, and high-proper-motion stars (HPM). 
Though most objects in the transient object catalogue belong to a single class, 
there is some uncertainty in the categorization of some of them.  
In this case, an interrogation sign is used when a class is not clear 
e.g. SN? or sometimes multiple possible classes are found for a single 
event e.g. SN/CV. 
Table \ref{table:top_classes} summarizes the number of objects in each class.

To compile the non-transient lightcurves we retrieve sources in the
dataset from the CRTS online catalogue by retrieving objects within a
20 arcsecond radius from CRTS detected transients, 
and removing known transient lightcurves from that set. 
We end up with 16940 unique non-transient lightcurves. Figure \ref{fig:non-transients} illustrates this process.

\begin{figure*}
\begin{center}
  \includegraphics[width=0.7\textwidth]{NonTransients.pdf}
\end{center}
  \caption{MANTRA Dataset Set Up: Lightcurve compilation for non-transients.}
  \label{fig:non-transients}
\end{figure*}

% Number of transients per transient class
\begin{table}
\centering
\begin{tabular}{c|c}
    \hline
    Class &  Object Count \\
    \hline
SN & 1723 \\
CV & 988 \\
HPM & 640 \\
AGN & 446 \\
SN? & 319 \\
Blazar & 243 \\
Unknown & 228 \\
Flare & 219 \\
AGN? & 138 \\
CV? & 77 \\
    \hline
\end{tabular}
\caption{Top 10 transient classes in the CRTS with their respective number of lightcurves.} 
\label{table:top_classes}
\end{table}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.45\textwidth]{cumulative_magnitude.pdf}
  \includegraphics[width=0.45\textwidth]{cumulative_classes.pdf}
\end{center}
  \caption{Cumulative number of lightcurves (expressed as a fraction)
    as a function of average magnitude (left) and number of data
    points in the lightcurve (right).
    This includes information for the three most representative
    classes (SN, CV, AGN) and the whole database (ALL).}
  \label{fig:cumulative}
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.6\textwidth]{examples_transient.pdf}
\end{center}
  \caption{Randomly selected lightcurves for the most represented transient classes as compiled in MANTRA. The class of each sample is within the legend box. }  
  \label{fig:examples_transient}
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.6\textwidth]{examples_nontransient.pdf}
\end{center}
  \caption{Randomly selected lightcurves for non-transient sources retrieved for MANTRA.}
  \label{fig:examples_non_transient}
\end{figure*}

Figure \ref{fig:cumulative} shows the number of lightcurves as a
function of average magnitude (left panel) and as a function of the
number of points in the lightcurve (right panel).
We show separately the whole data set and three most representative
classes: supernova, cataclysmic variables and active galactic nuclei. 
For these four sets, the median magnitude is in the range $18-20$. 
The number of points in the lightcurve has a larger variability.
The median for all the curves is close to 30, while for SN, CV and
AGN it is close to 15, 50 and 180, respectively. 
We provide sample lightcurves of the most represented transient classes 
and non-transient sources in Figure \ref{fig:examples_transient} and Figure \ref{fig:examples_non_transient}, respectively. 
The brightness evolution of non-transient sources is more stable over time, 
while transient objects present non-periodical changes at different time scales.

The challenges in the classification of lightcurves 
include the inherent nature of transient events, which is reflected 
in different brightness behaviors, their evolution over time, and 
the nonuniform sampling of observations at sequential dates. 
Besides, there is a large class imbalance to localize transient events, 
and perform their subsequent classification.

\subsection{Classification Tasks} \label{subsection_classification}
We study two classification tasks on the MANTRA dataset:

\begin{itemize}
\item {Binary Classification}.
Using a balanced number of events from both classes in order 
to investigate the capability of distinguishing between transient
and non-transient sources.
\item{8-Class Classification}.
Using the unbalanced number of objects across classes to 
perform a classification into the following categories:
AGN, Blazar, CV, Flare, HPM, Other, SN and Non-Transient.
\end{itemize}

We evaluate both tasks using the metrics of a detection problem. 
For each class in the testing set, we report the maximum F1-Score 
that is defined as the harmonic mean of precision and recall. 
We construct Precision-Recall (PR) curves by setting different 
thresholds on the output probabilities of belonging to each class.

\section{Repository Description} 
\label{sec:repository}

The repository contains the lightcurves and a Jupyter notebook
to reproduce some of the Figures and Tables in this paper.
The repository can be found in \url{https://github.com/MachineLearningUniandes/MANTRA}. 
To date the repository has two main folders:
\begin{itemize}

\item \texttt{data/lightcurves}: 
contains three text files in CSV format
the transient lightcurves (\texttt{transient\_lightcurves.csv}),
the labels for the transients (\texttt{transient\_labels.csv}) and
the lightcurves for non-transient objects
(\texttt{nontransient\_lightcurves.csv}). 
The first two files can be linked by unique transient IDs and
provided in the CRTS database. 
\item \texttt{nb-explore}: includes a jupyter notebook
  (\texttt{explore\_light\_curves.ipynb}) with examples on how to read
  and plot transient and non-transient lightcurves, extract the statistics in Table
  \ref{table:top_classes} and prepare the summary statistics in Figure
  \ref{fig:cumulative}. 
  Additional python files (\texttt{features.py},
  \texttt{helpers.py} and \texttt{inputs.py}) allow to read and perform
  simple operations on the CSV data files. 
\end{itemize}

\section{Machine Learning Methods} 
\label{sec:ml_tests}
Here we test baseline algorithms on the MANTRA dataset that can be
used as a reference for future work.
Figure \ref{fig:ML} shows an overview of our transient classification framework.
The main steps include data processing, feature extraction and classification.

\subsection{Preprocessing and Feature Extraction}

We do not input directly the annotated lightcurves to the ML algorithms.
We perform a preprocessing stage as follows. 
First, we discard lightcurves with less than 10 data points observations
as they may not contain enough information to be classified correctly.

\begin{figure*}
\begin{center}
  \includegraphics[width=0.9\textwidth]{ML.pdf}
\end{center}
  \caption{Overview of the Machine Learning process on the MANTRA
    dataset for the binary and 8-class classification tasks. We take
    the raw lightcurves as input, preprocess the data (step 1) and
    balance the classes for the training phase (step 2). 
    We extract the features (step 3) that are feed into the ML
    algorithms that perform the classification task (step 4).}
  \label{fig:ML}
\end{figure*}

Given that the number of lightcurves per class is imbalanced, 
in order to have the same number of instances for each class, we implement an
oversampling step by artificially generating multiple mock lightcurves 
from an observed one. 
We generate a slightly different lightcurve from the observed lightcurve and 
then sample the observed magnitude from a Gaussian distribution
centered on the observational apparent magnitude with the magnitude's
error as the standard deviation.

Finally, we compute a standard set of features for each lightcurve. 
These features are scalars derived from statistical and model-specific
fitting techniques. 
The first features (moment-based, magnitude-based and
percentile-based) were formally introduced in 
\cite{1101.1959}, and have been used in other studies \citep{1603.00882,1601.03931}.  
We extend that list to include another set (polynomial fitting-based features). 
At the end of this process we normalize the features to have zero
mean and unit variance.

These groups of features are:

\begin{enumerate}

\item Moment-based features, which use the magnitude for each lightcurve.
  \begin{itemize}
  \item \texttt{beyond1std}: 
    Percentage of observations which are over or under one standard
    deviation from the weighted average. Each weight is calculated as
    the inverse of the corresponding observation's photometric error. 
  \item \texttt{kurtosis}: 
    The fourth moment of the data distribution. 
  \item \texttt{skew}: 
    Skewness. Third moment of the data distribution.
  \item \texttt{sk}:
    Small sample kurtosis.
  \item \texttt{std}::
    The standard deviation.
  \item \texttt{stetson\_j}:
    The Welch-Stetson J variability index
    \citep{1996PASP..108..851S}. A robust standard deviation. 
  \item \texttt{stetson\_k}:  The Welch-Stetson K variability index
    \citep{1996PASP..108..851S}. A robust kurtosis measure. 
  \end{itemize}

\item Features based on the magnitudes.
    \begin{itemize}
    \item \texttt{amp}: 
      The difference between the maximum and minimum magnitudes.
    \item \texttt{max\_slope}: 
      Maximum absolute slope between two consecutive observations.
    \item \texttt{mad}: 
      The median of the difference between magnitudes and the median
      magnitude. 
    \item \texttt{mbrp}: 
      The percentage of points within 10\% of the median magnitude.
    \item \texttt{pst}: 
      Percentage of all pairs of consecutive magnitude measurements that have positive slope.
    \item \texttt{pst\_last30}: 
      Percentage of the last 30 pairs of consecutive magnitudes that
      have a positive slope, minus percentage of the last 30 pairs of
      consecutive magnitudes with a negative slope. 
    \end{itemize}

  \item Percentile-based features, which use the sorted flux distribution for
    each source. The flux is computed as $F = 10^{0.4 \mathrm{mag}}$. 
    We define $F_{n,m}$ as the difference between the $m$-th and $n$-the flux
    percentiles. 
    \begin{itemize}
    \item \texttt{p\_amp}: 
      Largest percentage difference between the absolute maximum
      magnitude and the median. 
    \item \texttt{pdfp}: 
      Ratio between $F_{5,95}$ and the median flux.
    \item \texttt{fpr20}: 
      Ratio $F_{40,60} / F_{5,95}$
    \item \texttt{fpr35}:
      Ratio $F_{32.5,67.5} / F_{5,95}$
    \item \texttt{fpr50}: 
      Ratio $F_{25,75} / F_{5,95}$
    \item \texttt{fpr65}: 
      Ratio $F_{17.5,82.5} / F_{5,95}$
    \item \texttt{fpr80}: 
      Ratio $F_{10,90} / F_{5,95}$
    \end{itemize}

  \item Polynomial Fitting-based features, which are the coefficients of
    multi-level terms in a polynomial curve fitting. This is a new set
    of features proposed in this paper. 
    \texttt{Polyn\_Tm} indicates the coefficient of the term of order
    \texttt{m} in a fit to a polynomial of order \texttt{n}.
    \begin{itemize}
        \item \texttt{Poly1\_T1}.
        \item \texttt{Poly2\_T1}.
        \item \texttt{Poly2\_T2}.
        \item \texttt{Poly3\_T1}.
        \item \texttt{Poly3\_T2}.
        \item \texttt{Poly3\_T3}.
        \item \texttt{Poly4\_T1}.
        \item \texttt{Poly4\_T2}.
        \item \texttt{Poly4\_T3}.
        \item \texttt{Poly4\_T4}.
    \end{itemize}    
\end{enumerate}

\subsection{ML algorithms}

We conduct experiments with three widely used families of supervised classification 
algorithms \citep{skysurvey, 1601.03931}: Neural Networks (NNs),
Random Forests (RFs) and Support Vector Machines (SVMs).

These algorithms are popular in published studies and are efficient 
for low dimensional feature datasets as is our case. 
We use sklearn \citep{1201.0490} Python's implementation of random
forests and support vector machines.  
Details on the inner workings of these machine learning models can be
found in \cite{9780387848570}.

We use the pytorch library for python for the development of the
linear Neural Networks. 
It consists of a series of fully connected layers that map the
features to the corresponding number of classes. At each layer, a 1d
batch normalization is implemented followed by a relu activtion
function. The final layer calls a softmax activation function to
transform the numerical values to class probabilities.

The hyperparameters explored for each algorithm are the
following.

\begin{itemize}
\item Neural Networks:
\begin{itemize}
\item Learning Rate: $\{0.1,0.01,0.001,0.0001\}$
\item Hidden Layer Sizes: Single , double or triple layers with $500$ nodes each.
\end{itemize}

\item Random Forest:
\begin{itemize}
    \item Number of Estimators: $200$ or $700$.
    \item Number of features considered: Square root or $log_2$ of the
      total number of features. 
\end{itemize}

\item Support Vector Machines:
\begin{itemize}
    \item Kernels: Radial Basis Function (RBF), linear or sigmoid.
    \item Kernel Coefficient ($\gamma$):  
      $\{0.125, 2, 32\}$
    \item Error Penalty (\textit{C}): $\{0.125, 2, 32\}$
\end{itemize}
\end{itemize}

\subsection{Validation} \label{subsection_importances}

We split the input lightcurves into training and testing in a $75:25$
ratio respectively, class by class. 
For the random forests and the SVM, we use a grid search over the
hyperparameter combinations with a 2-fold cross-validation over the
training set to determine the best hyperparameters.  
For the neural networks, at each epoch, the network is evaluated on
the test data.

\subsection{Results}

Table \ref{table:all-avg-results} shows the average class precision,
recall and F1-measure for each of the classification tasks and
algorithms liste above.

\begin{table}
\centering
% \resizebox{0.45\textwidth}{!}{%
\begin{tabular}{ccccc}
\hline
\multicolumn{1}{l}{\textbf{Case}} & \textbf{Classifier} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\ \hline \hline
\multirow{3}{*}{Binary}                 & RF                  & \textbf{86.61}    & \textbf{86.61} & \textbf{86.61}   \\
                                        & SVM                 & 82.61             & 82.525          & 82.57             \\
                                        & NN                  & 71.84              & 73.19           & 72.51             \\ \hline
\multirow{3}{*}{8 Class}                & RF                  & \textbf{46.25}     & \textbf{63.59}  & \textbf{50.38}             \\
                                        & SVM                 & 32.94              & 55.22           & 36.60             \\
                                        & NN                  & 36.0     & 55.19           & 38.59   \\ \hline
\end{tabular}%
% }
\caption{Average precision, recall and F1-score accross all classes
  for each algorithm and classification task. Best results per metric
  per classification task are in bold.} 
\label{table:all-avg-results}
\end{table}

\begin{figure}
\begin{center}
  \includegraphics[width=0.5\textwidth]{normalizedCMBinary.pdf}
\end{center}
  \caption{Confusion Matrix for the best performing model in the
    Binary task. Rows represent prediction and columns the ground
    truth.} 
  \label{fig:normalizedBinaryCM}
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{normalizedCM.pdf}
\end{center}
  \caption{Confusion Matrix for the best performing model in the 8-class
  task. The classes follow the abbreviations in Table
  \ref{Overall-Scores-8-Class-Regular}. Rows represent the
  predictions, and columns the ground truth.} 
  \label{fig:normalized8ClassCM}
\end{figure*}

%%%%%%  BINARY  %%%%%%
\subsubsection{Binary Classification} 
\label{Results-Binary}

The best algorithm in this task is RFs with an average F1-Score of
$86.61\%$.   
SVMs are the second best-performing model with a F1-Score of $82.57\%$. 
Changing the number of features does not significantly affect the score.
NNs are ranked third with an F1-Score of $72.51\%$

Figure \ref{fig:normalizedBinaryCM} shows the confusion matrix of the best
performing algorithm. These results suggest that in an imbalanced set
up, non-transient sources are better classified while transients are
more difficult, showing a difference of about  15 points in the
percentage of correct classifications.  
This difference in performance could be attributed to the intra-class
variation within the transients.

%%%%%%  EIGHT-CLASS  %%%%%%
\subsubsection{Eight-Class Classification}

For this task, RF is again the best classifier.
The best F1-Score is $50.38\%$. 
NNs are the second best model with an F1-Score is $36.60\%$, while
SVMs are the worst-performing model only achieving an average F1-Score
of $38.59\%$.   
Table \ref{Overall-Scores-8-Class-Regular} summarizes the results for 
individual classes and Figure  \ref{fig:normalized8ClassCM} presents 
the confusion matrix for the RF.

The two classes with highest F1-Score are non-transient ($87.12\%$) and
CV ($68.77\%$).  
The recall decreases for the non-transient class in comparison to the binary experiment, 
meaning that the algorithm misclassified some instances that belong to non-transient class
among transient classes. 
However, transient sources are not commonly confused with non-transient ones. 
The worst performing classes are Flare, Other and HPM, with F1-Scores in the 
range 11\% - 40\%. 
It is worth noting that the less frequent classes present a lower performance, 
such as Flare and HPM. 
Even though the most frequent classes are more easily identified, 
the "other" type class has a low F1-score due to the diverse nature 
of sources assigned to this category.

\begin{table}
\centering
% \resizebox{\textwidth}{!}{%
\begin{tabular}{ccccc}
\hline
\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score}& \textbf{Cover} \\\hline \hline
SN             & 48.82	& 51.39 &	50.07	 & 323 \\\hline
CV             & 66.96	& 70.69 &	68.77	 & 215 \\\hline
AGN            & 48.14	& 85.84 &	61.69	 & 106 \\\hline
HPM            & 25.19	& 86.84 &	39.05	 & 76 \\\hline
Blazar         & 46.77	& 49.15 &	47.93	 & 59 \\\hline
Flare          & 7.00	& 41.17 &	11.96	 & 51 \\\hline
Other          & 31.11	& 44.01 &	36.46	 & 234 \\\hline
Non-Tr.        & 96.06	& 79.69 &	87.12	 & 3798 \\\hline
avg/total      & 46.25	& 63.60 &	50.38	 & 4862 \\\hline
\end{tabular}%
\caption{Precision, Recall and F1-Score for the 8-Class Classification Task.}
\label{Overall-Scores-8-Class-Regular}
% }
\end{table}

\section{Conclusions}
\label{sec:conclusions}

The scope of forthcoming large astronomical synoptic surveys 
motivates the development and exploration of automatized ways to
detect transient sources. 
Such developments require observational datasets to train and test new
algorithms.  
Making these datasets public and easy to access has the potential
to open the field to a larger number of contributors.
With these goals in mind, in this paper we presented a compilation
based on data from the Catalina Real-Time Transient Survey.
The dataset has  $4869$ transient and $16940$ non-transient
lightcurves. 
The dataset is publicly available at
\url{https://github.com/MachineLearningUniandes/MANTRA}.

We illustrated how to use this database by extracting 
characteristic features to use them as input to train three different
machine learning algorithms (Random Forests, Neural Networks and
Support Vector Machines) for classification tasks.
The features extracted from lightcurves were either statistical
descriptors of the observations, or polynomial curve fitting
coefficients applied to the lightcurves.   
Overall, the best classifier for all tasks was the Random Forest.
Neural Networks showed the worst performance in the binary
classification task and the second best in the 8-class classification
task.

Certainly other classification algorithms could be used on the MANTRA
lightcurves. Our purpose here is not a thorough analysis of machine
learning algorithms. Our focus is in the compilation of the data into
a format that could easily be used for different projects that could
be addressed with MANTRA. 
For instance the supernovae detection problem could be studied as a
classfication problem between SN and the other classes, including
non-transients. 
Considering extremely unbalanced classes could also be addressed
with the dataset we are presenting by simply using less samples for
the SN class and keeping a large numbers for the non-transient class.

In a second paper we will present the second part of MANTRA.
It corresponds to almost one million images from the CRTS.
These will be tested using state-of-the art deep learning techniques
for transient classification.

\section*{Acknowledgements}

We thank Andrew Drake for sharing with us the CRTS Transient dataset
used in this project.  
We acknowledge funding from Universidad de los Andes in the call for
project finalization.
We also thank contributors and collaborators of the SciKit-Learn,
Jupiter Notebooks and Pandas' Python libraries.

CRTS and CSDR2 are supported by the U.S.~National Science 
Foundation under grant NSF grants AST-1313422, AST-1413600, and 
AST-1518308.  The CSS survey is funded by the National Aeronautics
and Space Administration under Grant No. NNG05GF22G issued through
the Science Mission Directorate Near-Earth Objects Observations Program.

### 2019-06-11 paper/main.tex
caption

@@
 \end{center}
-\caption{A}
+\caption{Distributions of peculiar velocities for MW type halos
+  (individual, paired and observed). 
+  Left panel shows the estimated probability density.
+  The right panel shows the cumulative probability density.
+  Paired halos have a slightly lower peculiar velocity than isolated
+  halos.
+  The observed peculiar velocity ($v_p=627\pm22$ km s$^{-1}$) shown as
+  the vertical shaded stripe, is always located on the high velocity
+  tail of simulations.}
 \label{fig:proba}

### 2019-06-11 code/study_pairs_abacus.py
update study pairs abacus

@@ -42,144 +42,36 @@ def load_box(BoxID=0):
     return datos
         
 
-def compute_pairs_FOF_abacus_box(BoxID=0, v_limit=627):
+def compute_pairs_FOF_abacus_box(BoxID=0):
     pair_count = {}
-    datos = load_box(BoxID=BoxID)       
-    
-    fig = plt.figure(figsize=(16,24))
-
-    # Input sample
-    plt.subplot(4,3,1)
-    _ = plt.hist(datos['vel_AB_rad'], bins=np.linspace(-500,300, 30))
-    plt.xlabel("Radial Velocity [km/s]")
-
-    plt.subplot(4,3,2)
-    plt.plot(np.sort(datos['vmax_G']), np.linspace(0,1,len(datos['vmax_G'])))
-    plt.plot(np.sort(datos['vmax_A']), np.linspace(0,1,len(datos['vmax_A'])))
-    plt.plot(np.sort(datos['vmax_B']), np.linspace(0,1,len(datos['vmax_B'])))
-    plt.xlabel("Vmax [km/s]")
-    plt.xlim(200,800)
-    plt.ylim(0,1)
-
-    plt.subplot(4,3,3)
-    plt.plot(np.sort(datos['vel_G_mag']), np.linspace(0,1,len(datos['vel_G_mag'])), color='black')
-    plt.plot(np.sort(datos['vel_A_mag']), np.linspace(0,1,len(datos['vel_A_mag'])))
-    plt.plot(np.sort(datos['vel_B_mag']), np.linspace(0,1,len(datos['vel_B_mag'])))
-    plt.xlim(0,1000)
-    plt.ylim(0,1)
-    plt.vlines(v_limit, 0,1)
-    plt.xlabel("Peculiar Velocity [km/s]")
-    plt.grid()
-    plt.title('Npairs={}, '.format(len(datos['vel_AB_rad']))+' Nhalos={}'.format(len(datos['vmax_G'])))
-    pair_count['total_pair'] = len(datos['vmax_A'])
-    pair_count['total_halos'] = len(datos['vmax_G'])
+    data = load_box(BoxID=BoxID)
+    print("processing box ", BoxID)
+    vlim = {'mean':627, 'sigma':22}
 
     # Selection in vmax
-    ii = ((datos['vmax_A'] < 240) & (datos['vmax_B']<240)) 
-    ll = (datos['vmax_G'] < 240)
-
-    plt.subplot(4,3,4)
[…]

### 2019-06-13 code/study_pairs_abacus.py
alingnment

@@ -30,10 +30,12 @@ def load_box(BoxID=0):
     #now we compute the radial velocity including the hubble flow
     data['vel_AB_rad'] = data['vel_AB_rad'] + (data['pos_AB'] * hubble)
 
+    # here we compute the compute the dot product between the position vector and the radial velocity
+    data['mu'] = np.abs(np.sum((data['pos_B']-data['pos_A'])*(data['vel_B']), axis=1)/(data['vel_B_mag']*data['pos_AB']))
 
     datos = {}
     ii = (data['pos_A'][:,0] > 10) & (data['pos_A'][:,0]<710)
-    keys = ['vel_A_mag', 'vel_B_mag', 'pos_AB', 'vel_AB', 'vel_AB_rad', 'vel_AB_tan', 'vmax_A', 'vmax_B']
+    keys = ['vel_A_mag', 'vel_B_mag', 'pos_AB', 'vel_AB', 'vel_AB_rad', 'vel_AB_tan', 'vmax_A', 'vmax_B', 'mu']
     for kk in keys:
         datos[kk] = data[kk][ii]
     keys = ['vel_G_mag', 'vmax_G']
@@ -42,36 +44,43 @@ def load_box(BoxID=0):
     return datos
         
 
-def compute_pairs_FOF_abacus_box(BoxID=0):
+def count_pairs_FOF_abacus_box(BoxID=0):
     pair_count = {}
     data = load_box(BoxID=BoxID)
     print("processing box ", BoxID)
     vlim = {'mean':627, 'sigma':22}
 
     # Selection in vmax
-    ii = (data['vmax_A']<240) & (data['vmax_B']<240)
+    ii = (data['vmax_A']<240) & (data['vmax_B']<240) 
+    ii &= (data['vel_AB_rad']<0) 
+    ii &= (np.abs(data['vel_AB_rad'])>data['vel_AB_tan'])
+    
     ll = (data['vmax_G']<240)
     
     # Selection of high velocity
-    jj = (data['vel_A_mag'] > vlim['mean'])
+    jj = (data['vel_B_mag'] > vlim['mean'])
     mm = (data['vel_G_mag'] > vlim['mean'])
     
     # Count
     pair_count['pair_total'] = np.count_nonzero(ii)
-    pair_count['individual_total'] = np.count_nonzero(ll)
     pair_count['pair_high'] = np.count_nonzero(ii&jj)
+    pair_count['individual_total'] = np.count_nonzero(ll)
[…]

### 2019-06-13 code/study_pairs_abacus.py
compile std

@@ -66,8 +66,10 @@ def count_pairs_FOF_abacus_box(BoxID=0):
     pair_count['pair_high'] = np.count_nonzero(ii&jj)
     pair_count['individual_total'] = np.count_nonzero(ll)
     pair_count['individual_high'] = np.count_nonzero(ll&mm)
-    pair_count['mean_mu_total'] = np.mean(np.mean(data['mu'][ii]))
-    pair_count['mean_mu_high'] = np.mean(np.mean(data['mu'][ii&jj]))
+    pair_count['mean_mu_total'] = np.median(data['mu'][ii])
+    pair_count['mean_mu_high'] = np.median(data['mu'][ii&jj])
+    pair_count['std_mu_total'] = np.std(data['mu'][ii])
+    pair_count['std_mu_high'] = np.std(data['mu'][ii&jj])
     return pair_count
 
 def all_data():
@@ -75,12 +77,14 @@ def all_data():
     f = open(outfile, "w")
     for i in range(40):
         pair_count = count_pairs_FOF_abacus_box(BoxID=i)
-        f.write("{:d} {:d} {:d} {:d} {:d} {:f} {:f}\n".format(i, 
+        f.write("{:d} {:d} {:d} {:d} {:d} {:f} {:f} {:f} {:f}\n".format(i, 
                                                    pair_count['individual_total'], 
                                                    pair_count['individual_high'],
                                                    pair_count['pair_total'],
                                                    pair_count['pair_high'],
                                                    pair_count['mean_mu_total'],
-                                                   pair_count['mean_mu_high']))
+                                                   pair_count['mean_mu_high'],
+                                                   pair_count['std_mu_total'],
+                                                   pair_count['std_mu_high']))
     f.close()
     

### 2019-06-25 paper/main.tex
new plots in main
new plot
edit

@@
-\documentclass[twocolumn]{aastex62}
+\documentclass[onecolumn]{aastex62}
 \submitjournal{ApJL}
@@
 \label{fig:proba}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=0.4\textwidth]{cumulative_alignment_06.pdf}
+\end{center}
+\caption{Cumulative probability for the dot product between the unit
+  vector joining the two dominan halos in the pair and the peculiar
+  velocity of the less massive halo.
+  Continous line: results for all pairs; dashed line: pairs where the
+  less massive halo moves with a pecular velocity larger than
+  $v_p=627\pm22$ km s$^{-1}$.}
+\label{fig:alignment}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.0\textwidth]{ratio_pair_total_ind_total.pdf}
+\end{center}
+\caption{Ratio}
+\label{fig:ratio_total}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.0\textwidth]{ratio_pair_high_ind_total.pdf}
+\end{center}
+\caption{Ratio}
+\label{fig:ratio_fast}
 \end{figure*}

### 2019-06-25 paper/main.tex
peculiar velocity

@@
   the vertical shaded stripe, is always located on the high velocity
-  tail of simulations.}
+  tail of simulations.
+  Halos in isolated pairs tend to have lower peculiar velocities than
+  the general sample of individual halos.}
 \label{fig:proba}
@@
 \caption{Cumulative probability for the dot product between the unit
-  vector joining the two dominan halos in the pair and the peculiar
+  vector joining the two halos in the pair and the peculiar
   velocity of the less massive halo.
@@
   less massive halo moves with a pecular velocity larger than
-  $v_p=627\pm22$ km s$^{-1}$.}
+  $v_p=627\pm22$ km s$^{-1}$.
+  The vertical stripe shows the values for the MW and M31.
+  Halos with high peculiar velocities tend to be aligned in the
+  direction towards its partner.}
 \label{fig:alignment}

### 2019-06-26 paper/main.tex
edit

@@
 \end{center}
-\caption{Ratio}
+\caption{Ratio of the number of halos in pairs to the total number of isolated
+  halos as a function of the cosmological parameters in the abacus
+  simulations. Across simulations only $0.9 (\pm 0.2) \%$ of the halos can
+  be found in pairs with kinematics and isolation similar to the
+  MW-M31 pair.} 
 \label{fig:ratio_total}
@@
 \end{center}
-\caption{Ratio}
+\caption{Ratio of the number of fast halos in pairs to the total
+  number of isolated halos as a function of the cosmological
+  parameters in the abacus simulations. Only $0.07 (\pm0.02) \%$ of
+  the halos have similar kinematic conditions as the MW-M31 pairs.}
 \label{fig:ratio_fast}

### 2019-07-02 paper/main.tex
edit

@@
 \submitjournal{ApJL}
+\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
+\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc}\fi}}  
+\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
+\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc}\fi}}  
+\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
+\newcommand{\kms}{{\ifmmode{{\rm km\, {s}^{-1}}}\else{km s$^{-1}$}\fi}}
+
 \title{Our peculiar motion across the Universe}
@@
 
+\section{Numerical Setup}
+We use simulations from the Abacus project \cite{abacus}.
+
+We use the simulation boxes with size $L_{\rm box}=720$\hMpc simulated
+with $1440^3$ particles and a resulting particle mass of $\sim 1\times
+10^{10}$\hMpc. 
+We use the set of 41 boxes with different $\omega$CDM cosmological
+parameters which are phase-matched to isolate cosmology dependences.
+
+\section{Sample Selection}
+
+We select all halos with a maximum circular velocity larger than $xx$\kms.
+
+\section{Results}
+
+\section{Conclusions}
+
 \cite{2009PhRvD..80d3005E}
@@
 \begin{center}
-  \includegraphics[width=0.4\textwidth]{cumulative_alignment_06.pdf}
-\end{center}
-\caption{Cumulative probability for the dot product between the unit
-  vector joining the two halos in the pair and the peculiar
-  velocity of the less massive halo.
-  Continous line: results for all pairs; dashed line: pairs where the
-  less massive halo moves with a pecular velocity larger than
-  $v_p=627\pm22$ km s$^{-1}$.
-  The vertical stripe shows the values for the MW and M31.
-  Halos with high peculiar velocities tend to be aligned in the
-  direction towards its partner.}
-\label{fig:alignment}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
   \includegraphics[width=1.0\textwidth]{ratio_pair_total_ind_total.pdf}
@@
 \label{fig:ratio_fast}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=0.4\textwidth]{cumulative_alignment_06.pdf}
+\end{center}
+\caption{Cumulative probability for the dot product between the unit
+  vector joining the two halos in the pair and the peculiar
+  velocity of the less massive halo.
+  Continous line: results for all pairs; dashed line: pairs where the
+  less massive halo moves with a pecular velocity larger than
+  $v_p=627\pm22$ km s$^{-1}$.
+  The vertical stripe shows the values for the MW and M31.
+  Halos with high peculiar velocities tend to be aligned in the
+  direction towards its partner.}
+\label{fig:alignment}
 \end{figure*}

### 2019-07-02 paper/main.tex
edit

@@
 
-We select all halos with a maximum circular velocity larger than $xx$\kms.
+We start by selectin all halos with a maximum circular velocity larger
+than $200$ \kms.
+Then we proceed to use these halos to find what we call an Isolated Pair
+(IP).
+IPs are two halos, $A$ and $B$, that do not have any other third halo
+closer than the $AB$ separation; in other words in IPs the two halos
+are mutually its closest neighbor.
 

### 2019-07-03 paper/main.tex
edit

--- 2019-06-26 paper/main.tex
@@
 \submitjournal{ApJL}
+\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc\ }\fi}}  
+\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc\ }\fi}}  
+\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc\ }\fi}}  
+\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc\ }\fi}}  
+\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
+\newcommand{\kms}{{\ifmmode{{\rm km\, {s}^{-1}}}\else{km s$^{-1}$}\fi}}
+
 \title{Our peculiar motion across the Universe}
@@
 
+\section{Numerical Setup}
+
+We use simulations boxes from the Abacus project \cite{abacus}. 
+Each box has a size of $L_{\rm box}=720$\hMpc and was simulated
+with $1440^3$ particles and a resulting particle mass of $\sim 1\times
+10^{10}$\hMpc.
+Each box also have 41 different $\omega$CDM cosmological parameters
+which are phase-matched to isolate cosmology dependences.
+In our analysis we use the Friends-of-Friends (FOF) catalogs built on
+the snapshot at redshift of $z=0.1$.
+
+\section{Sample Selection}
+
+We start by selecting all halos with a maximum circular velocity,
+in the range $200\kms < v_{\rm max}< 250 \kms$ 
+Then we proceed to use these halos to find what we call an Isolated Pair
+(IP).
+IPs are two halos, $A$ and $B$, that do not have any other third halo
+with $v_{\rm max}>200$\kms closer than the pair separation.
+We follow a convention where $A$ refers to the least massive halo in
+the pair.
+
+From this pair we measure the following quantities: the maximum
+circular velocity for the two halos, peculiar velocities $\vec{v}_p$, 
+separation vector $\vec{r}_{AB}$, tangential velocity of $B$ with
+respect to $A$, $v_{t}$ and the radial velocity of $B$ with respect
+to $B$ after taking into account Hubble expansion $v_r$.
+
+From the IP sample we select the pairs with some similarity to the
+observed LG. 
+The conditions we impose are the following: the maximum
+circular velocity should for both halos in the pair must be less than
+$250$ \kms, the relative radial velocity should be negative and the
+magnitude of the tangential velocity should be lower than the
+magnitude of the radial velocity. 
+We refer to this sample simply as the pair sample. 
+Only about $1\%$ of the halos in the initial sample are found in
+the pair sample. 
+This is consistent with previous results that show that the fraction
+of LG pairs in a cosmological context
+
+\section{Results}
+
+\section{Conclusions}
+
 \cite{2009PhRvD..80d3005E}
